---
title: "MB"
author: "Isolde Lane Shaw"
date: "1/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=3)
```


# MAPPING THE 1D AND 2D POST-HOC BINNING PREDICTIONS 

Here I create rasters showing the birdPreds predictions according to the classes given in the landscape rasters. 

Note: can take significant memory

```{r files needed mapBins, include=FALSE}

# These rasters will be needed for this section
load(file.path(downloadFolderForestClass, "forClassRaster.Rdata"))
load(file.path(downloadFolderForestClass, "nonForRaster.Rdata"))
load(file.path(downloadFolderForestClass, "ageRaster.Rdata"))
load(file.path(outputFolderBirdPreds, "birdPreds.Rdata"))

#I will need a birdList that includes only birds included in the birdPreds
birdList <- birdList

```
 
 
# 1D mapBins

### 1D mapBins - nonForRaster area only

```{r 1D mapBins nonForested}

#get non-Forest 1D data together
nf1DPreds <- lapply(X = birdList, FUN = function(bird) {

#separate out data table rows that are forested, get rid of unnecessary forestedStatus column
nonforestedDF <- eval(parse(text=paste("birdPreds$birdPreds1D$", bird, sep = "")))  
nonforestedDF <- nonforestedDF[forestedStatus == "0"]
nonforestedDF <- nonforestedDF[ , c(3,5)]
nonforestedDF <- droplevels(nonforestedDF)

return(nonforestedDF)
})

names(nf1DPreds) <- birdList
nf1DPreds

#reclassify non forest raster to get map of 1D bird preds in non forested areas
nf1DMaps <- lapply(X = birdList, FUN = function(bird){


  nfBirdPreds <- eval(parse(text=paste("nf1DPreds$", bird, sep = "")))
      
#make numeric
nfBirdPreds <- nfBirdPreds[, landForClass:=as.numeric(landForClass)]
nfBirdPreds <- nfBirdPreds[, meanBirdDensity:=as.numeric(meanBirdDensity)]
str(nfBirdPreds) #check

#raster1DBins <- nonForRaster 
raster1DBins <- reclassify(nonForRaster, nfBirdPreds)

names(raster1DBins) <- paste(bird)
plot(raster1DBins)

print(paste(bird,"nf 1D map raster complete"))
return(raster1DBins)
    })

  names(nf1DMaps) <- birdList


```

```{r save 1Dbins nonFor rasters}

#as Rdata file

   save(nf1DMaps,
        file =  file.path(outputFolderBirdPredsRasters, "nf1DMaps.Rdata"))
   # #load(file.path(outputFolderBirdPredsRasters, "nf1DMaps.Rdata"))

#as tif files

# 
#  lapply(X = birdList, FUN = function(bird){
#  
#   raster <- eval(parse(text=paste("rasters1DBinsNF$", bird, sep = "")))
#   names(raster) <- paste(bird)
#   writeRaster(x = raster, filename = file.path(outputFolderBirdPredsRasters, 
#                                                 paste(bird, "-1D-NF", sep = "")), 
#                                                 format = "GTiff", 
#                                                 overwrite = TRUE)
#    
#     })


```


## 1D mapBins - forClassRaster area only

```{r 1D for maps}

#get non-Forest 1D data together
for1DPreds <- lapply(X = birdList, FUN = function(bird) {

#separate out data table rows that are forested, get rid of unnecessary forestedStatus column
forestedDF <- eval(parse(text=paste("birdPreds$birdPreds1D$", bird, sep = "")))  
forestedDF <- forestedDF[forestedStatus == "1"]
forestedDF  <- forestedDF [ , c(3,5)]
forestedDF <- droplevels(forestedDF)

return(forestedDF)
})

names(for1DPreds) <- birdList
for1DPreds

#reclassify forest class raster to give 1D bird prediction values for each bird sp
for1DMaps <- lapply(X = birdList, FUN = function(bird){


nfBirdPreds <- eval(parse(text=paste("for1DPreds$", bird, sep = "")))
      
#make numeric
nfBirdPreds <- nfBirdPreds[, landForClass:=as.numeric(landForClass)]
nfBirdPreds <- nfBirdPreds[, meanBirdDensity:=as.numeric(meanBirdDensity)]
str(nfBirdPreds) #check

#raster1DBins <- nonForRaster 
raster1DBinsForest <- reclassify(forClassRaster, nfBirdPreds)

names(raster1DBinsForest) <- paste(bird)
plot(raster1DBinsForest)


print(paste(bird,"for 1D map raster complete"))
return(raster1DBinsForest)
    })

  names(for1DMaps) <- birdList


```

```{r save 1Dbins forest rasters}

#as Rdata file
save(for1DMaps,
     file =  file.path(outputFolderBirdPredsRasters, "for1DMaps.Rdata"))
#load(file.path(outputFolderBirdPredsRasters, "for1DMaps.Rdata"))

#as tif files


 # lapply(X = birdList, FUN = function(bird){
 # 
 #  raster <- eval(parse(text=paste("rasters1DBinsForest$", bird, sep = "")))
 #  names(raster) <- paste(bird)
 #  writeRaster(x = raster, filename = file.path(outputFolderBirdPredsRasters, 
 #                                                paste(bird, "-1D-for", sep = "")), 
 #                                                format = "GTiff", 
 #                                                overwrite = TRUE)
 #   
 #    })


```


### Create composite map of 1D predictions for forClassraster areas and 1D predictions for nonForRaster areas

```{r combine 1D for and nonfor rasters}


for1DAndNf1DMaps <- lapply(X = birdList, FUN = function(bird){
 
  raster1DBinsNF <- eval(parse(text=paste("nf1DMaps$", bird, sep = "")))
  raster2DBinsFor <- eval(parse(text=paste("for1DMaps$", bird, sep = "")))
  
birdPredsRaster1D <- cover(x = raster2DBinsFor,
                         y = raster1DBinsNF)
                        
names(birdPredsRaster1D) <- paste(bird)

birdPredsRaster1D #visually check Raster
plot(birdPredsRaster1D)



print(paste(bird,"for 1D and nf 1D map complete"))
return(birdPredsRaster1D)
    })

names(for1DAndNf1DMaps) <- birdList

```

```{r save for1DAndNf1DMaps}

#as Rdata file
   save(for1DAndNf1DMaps,
        file =  file.path(outputFolderBirdPredsRasters, "for1DAndNf1DMaps.Rdata"))
   #load(file.path(outputFolderBirdPredsRasters, "for1DAndNf1DMaps.Rdata"))

#as tif files

# 
#  lapply(X = birdList, FUN = function(bird){
#  
#   raster <- eval(parse(text=paste("birdPredsRasters1D$", bird, sep = "")))
#   names(raster) <- paste(bird)
#   writeRaster(x = raster, filename = file.path(outputFolderBirdPredsRasters, 
#                                                 paste(bird, "-1D-birdPreds", sep = "")), 
#                                                 format = "GTiff", 
#                                                 overwrite = TRUE)
#    
#     })


```



# 2D mapBins  

Here I create a rasters of bird density using the 2D binning predictions for forested areas. 


## Reclassify the ageRaster to ageClassRaster 

First, the age raster needs to be reclassified into the age classes specified in the ageClassDefs table included in the birdPreds. 


```{r make ageClassRaster}

#reclassify forAgeRaster into a raster of forest age classes
ageReClassTab <- birdPreds$ageClassDefs
ageReClassTab <- ageReClassTab[ , ageClasses:=as.numeric(ageClasses)] #change data type of ageClassDefs
str(ageReClassTab) #check
ageClassRaster <- ageRaster #make copy of forAgeRaster to be reclassified
ageClassRaster <- reclassify(ageClassRaster, ageReClassTab) #do the reclassification based on ageClassDefs
names(ageClassRaster) <- "ageClassRaster" 
ageClassRaster#check over the raster that has been reclassified
plot(ageClassRaster)

```


### 2D mapBins - forClassRaster area only

```{r for 2D maps }

for2DMaps <- lapply(X = birdList, FUN = function(bird){

# check that spatial extent is the same for ageClassraster and forClassraster
extent(forClassRaster) == extent(ageClassRaster)

#reform matrix
matrix <- eval(parse(text=paste("birdPreds$birdMatricies$", bird, sep = "")))
reclassTab2D <- melt(matrix)
colnames(reclassTab2D) <- c( "forClass","ageClass", "birdDensityPred")

#reclassify Raster according to reclassTab2D, ageClassRaster and forClassRaster
raster2DBins <- raster(forClassRaster); raster2DBins[] = NA #make an empty NA raster the same as forClassRaster

#make dataframe of all the data in forClassRaster and ageClassRaster and give each cell/row a new definition column, birdDensityPred, from reclassTab2d
f = data.frame(forClass=forClassRaster[], ageClass=ageClassRaster[])
vec = c(1:nrow(f))
f[,3] = vec
m = merge(f, reclassTab2D, all.x=TRUE)
colnames(m)[3] = "ord"
m = m[order(m$ord),]
#populate raster2DBins with the birdDensityPred row of the table m
raster2DBins[] = m$birdDensityPred

names(raster2DBins) <- paste(bird)

#check the new raster
raster2DBins
plot(raster2DBins) 

 print(paste(bird,"for 2D map raster complete"))

      return(raster2DBins)
    })

names(for2DMaps) <- birdList

```

```{r save for 2D maps }

#as Rdata file

   save(for2DMaps,
        file =  file.path(outputFolderBirdPredsRasters, "for2DMaps.Rdata"))
   # #load(file.path(outputFolderBirdPredsRasters, "for2DMaps.Rdata"))

#as tif files

# 
#  lapply(X = birdList, FUN = function(bird){
#  
#   raster <- eval(parse(text=paste("rasters2DBins$", bird, sep = "")))
#   names(raster) <- paste(bird)
#   writeRaster(x = raster, filename = file.path(outputFolderBirdPredsRasters, 
#                                                 paste(bird, "-2D-ForestRaster", sep = "")), 
#                                                 format = "GTiff", 
#                                                 overwrite = TRUE)
#    
#     })


```


### Create composite map of 2D predictions for forClassraster areas and 1D predictions for nonForRaster areas

```{r combine 1D and 2D rasters}


for2DAndNf1DMaps <- lapply(X = birdList, FUN = function(bird){
  
  raster1DBins <- eval(parse(text=paste("nf1DMaps$", bird, sep = "")))
  raster2DBins <- eval(parse(text=paste("for2DMaps$", bird, sep = "")))
  
birdPredsRaster <- cover(x = raster2DBins,
                         y = raster1DBins) 

names(birdPredsRaster) <- paste(bird)

birdPredsRaster #visually check Raster
plot(birdPredsRaster)

# writeRaster(x = birdPredsRaster, filename = file.path(outputFolderBirdPredsRasters, paste(bird, "-birdPredsRaster", sep = "")), format = "GTiff", overwrite = TRUE)

print(paste(bird,"for 2D and nf 1D map raster complete"))
return(birdPredsRaster)
    })

names(for2DAndNf1DMaps) <- birdList

```

```{r save for2DAndNf1DMaps}

#as Rdata file

   save(for2DAndNf1DMaps,
        file =  file.path(outputFolderBirdPredsRasters, "for2DAndNf1DMaps.Rdata"))
   #load(file.path(outputFolderBirdPredsRasters, "for2DAndNf1DMaps.Rdata"))

#as tif files


 # lapply(X = birdList, FUN = function(bird){
 # 
 #  raster <- eval(parse(text=paste("birdPredsRasters$", bird, sep = "")))
 #  names(raster) <- paste(bird)
 #  writeRaster(x = raster, filename = file.path(outputFolderBirdPredsRasters, 
 #                                                paste(bird, "-birdPredsRaster2D", sep = "")), 
 #                                                format = "GTiff", 
 #                                                overwrite = TRUE)
 #   
 #    })
 # 

```




# CREATION OF RESIDUAL RASTERS

Here rasters are created that show the residual values between the original birdRasters and the mapBin rasters created from the post-hoc binning birdPreds predictions of bird density. 

Note: can take significant memory

```{r files needed create residual rasters, include=FALSE}

# These rasters will be needed for this section
load(file.path(downloadFolderBird, "birdRasters.Rdata"))
load(file.path(outputFolderBirdPredsRasters, "for1DMaps.Rdata"))
load(file.path(outputFolderBirdPredsRasters, "for2DMaps.Rdata"))
load(file.path(outputFolderBirdPredsRasters, "for1DAndNf1DMaps.Rdata"))
load(file.path(outputFolderBirdPredsRasters, "for2DAndNf1DMaps.Rdata"))

#I will need a birdList that includes only birds included in the above lists
birdList <- birdList

```


### Make 1D residual rasters - forClassRaster area only

```{r getResiduals forest 1D}



for1DRes <- lapply(X = birdList, FUN = function(bird){
  
  NM <- eval(parse(text=paste("birdRasters$", bird, sep = "")))
  MB <- eval(parse(text=paste("for1DMaps$", bird, sep = "")))
  
res <- overlay(NM, MB, fun=function(x,y){return(x-y)})

names(res) <- paste(bird)
plot(res)

print(paste(bird," forest 1D res raster complete"))
return(res)
    })

names(for1DRes) <- birdList

#as Rdata file

   save(for1DRes,
        file =  file.path(outputFolderBirdPredsRasters, "for1DRes.Rdata"))
   # load(file.path(outputFolderBirdPredsRasters, "for1DRes.Rdata"))
   
```


### Make 2D residual rasters - forClassRaster area only

```{r getResiduals forest 2D}



for2DRes <- lapply(X = birdList, FUN = function(bird){
  
  NM <- eval(parse(text=paste("birdRasters$", bird, sep = "")))
  MB <- eval(parse(text=paste("for2DMaps$", bird, sep = "")))
  
res <- overlay(NM, MB, fun=function(x,y){return(x-y)})

names(res) <- paste(bird)
plot(res)

print(paste(bird," forest 2D res raster complete"))
return(res)
    })

names(for2DRes) <- birdList

#as Rdata file

   save(for2DRes,
        file =  file.path(outputFolderBirdPredsRasters, "for2DRes.Rdata"))
   # load(file.path(outputFolderBirdPredsRasters, "for2DRes.Rdata"))
   

```



### Make residual rasters of composite 1D predictions for forClassraster areas and 1D predictions for nonForRaster areas

```{r getResiduals for 1D and nf 1D }



for1DAndNf1DRes <- lapply(X = birdList, FUN = function(bird){
  
  NM <- eval(parse(text=paste("birdRasters$", bird, sep = "")))
  MB <- eval(parse(text=paste("for1DAndNf1DMaps$", bird, sep = "")))
  
res <- overlay(NM, MB, fun=function(x,y){return(x-y)})

names(res) <- paste(bird)
plot(res)

print(paste(bird," for 1D and nf 1D res raster complete"))
return(res)
    })

names(for1DAndNf1DRes) <- birdList

#as Rdata file

   save(for1DAndNf1DRes,
        file =  file.path(outputFolderBirdPredsRasters, "for1DAndNf1DRes.Rdata"))
   # load(file.path(outputFolderBirdPredsRasters, "for1DAndNf1DRes.Rdata"))
   
```


### Make residual rasters of composite 2D predictions for forClassraster areas and 1D predictions for nonForRaster areas

```{r getResiduals for2D and nf1D}



for2DAndNf1DRes <- lapply(X = birdList, FUN = function(bird){
  
  NM <- eval(parse(text=paste("birdRasters$", bird, sep = "")))
  MB <- eval(parse(text=paste("for2DAndNf1DMaps$", bird, sep = "")))
  
res <- overlay(NM, MB, fun=function(x,y){return(x-y)})

names(res) <- paste(bird)
plot(res)

print(paste(bird," for2D and nf1D res raster complete"))
return(res)
    })

names(for2DAndNf1DRes) <- birdList

#as Rdata file

   save(for2DAndNf1DRes,
        file =  file.path(outputFolderBirdPredsRasters, "for2DAndNf1DRes.Rdata"))
   # load(file.path(outputFolderBirdPredsRasters, "for2DAndNf1DRes.Rdata"))
   
```




# VISUALISATION OF RASTERS



```{r plot NMs maps and Res for comparaison}

getResRasComparison <- function(birdName, phbRasterListName, resRasterListName) {

  natMod <- eval(parse(text=paste("birdRasters$", birdName, sep = "")))
  phbRas <- eval(parse(text=paste(phbRasterListName, "$", birdName, sep = "")))
  resRas <- eval(parse(text=paste(resRasterListName, "$", birdName, sep = "")))
 names(natMod) <- "National Model"
 names(phbRas) <- "Post-Hoc Binning"
 names(resRas) <- "Residuals"

compRas <- raster::stack(natMod, phbRas, resRas)


rasComparison <- gplot(compRas) + 
                  geom_tile(aes(fill = value)) +
                    facet_wrap(~ variable) +
                    scale_fill_gradientn(colours = rev(terrain.colors(500)), na.value = "grey70") +
  coord_equal() + 
    ggtitle(paste0(birdName)) +
     theme_classic() +
  theme(legend.position = "bottom",
        line = element_blank(),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        legend.key.width = unit(2, 'cm'))
  #scale_fill_gradientn(colours=c("royalblue4", "orangered1", "springgreen4", "violetred",  "aquamarine1", "goldenrod4", "#0000FFFF","#FFFFFFFF","#FF0000FF"), na.value = "grey70")

  return(rasComparison)
}


resRasComparison <- getResRasComparison(birdName = "BBWO", 
                                  phbRasterListName = "for2DAndNf1DMaps",
                                  resRasterListName = "for2DAndNf1DRes")

resRasComparison
```

```{r plot NM and maps for comparason}

getResRasComparison <- function(birdName, phbRaster1DListName, phbRaster2DListName) {

  natMod <- eval(parse(text=paste("birdRasters$", birdName, sep = "")))
  phbRas1D <- eval(parse(text=paste(phbRaster1DListName, "$", birdName, sep = "")))
    phbRas2D <- eval(parse(text=paste(phbRaster2DListName, "$", birdName, sep = "")))
  
 names(natMod) <- "National Model"
 names(phbRas1D) <- "Post-Hoc Binning 1D"
  names(phbRas2D) <- "Post-Hoc Binning 2D"
 
 
 

compRas <- raster::stack(natMod, phbRas1D, phbRas2D)


rasComparison <- gplot(compRas) + 
                  geom_tile(aes(fill = value)) +
                    facet_wrap(~ variable) +
                     scale_fill_gradientn(colours = rev(terrain.colors(500)), na.value = "grey70") +
    coord_equal() + 
    ggtitle(paste0(birdName)) +
     theme_classic() +
  theme(legend.position = "bottom",
        line = element_blank(),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        legend.key.width = unit(2, 'cm')) 
  #scale_fill_gradientn(colours=c( "white", "darkgreen"), na.value = "grey70")
      #scale_fill_viridis_c(breaks = quantile(rank(df$z)),
                       #labels = quantile(df$z))
#"blue", "lightblue",  "lightgreen", "green", "yellow", "orange", "orangered1", "red", "violetred"
  return(rasComparison)
}


resRasComparison <- getResRasComparison(birdName = "BARS", 
                                  phbRaster1DListName = "for1DAndNf1DMaps",
                                  phbRaster2DListName = "for2DAndNf1DMaps")

resRasComparison


# pal <- colorRampPalette(c("blue", "darkblue", "darkgreen", "lightgreen", "yellow", "orange", "red", "purple"))
# 
 plot(birdRasters$BARS, colNA="grey70")

 plot(for2DAndNf1DMaps$BARS, colNA="grey70")
 plot(for2DAndNf1DRes$BARS,  colNA="grey70")
 plot(for2DRes$OVEN,  colNA="grey70")
 plot(for1DRes$OVEN,  colNA="grey70")
# 
breaks_qt <- classInt::classIntervals(na.omit(getValues(birdRasters$OVEN)), n = 7, style = "quantile")
plot(birdRasters$OVEN, breaks=breaks_qt$brks, col = topo.colors(7), colNA="grey70")
breaks_qt <- classInt::classIntervals(na.omit(getValues(for2DAndNf1DMaps$OVEN)), n = 7, style = "quantile")
plot(for2DAndNf1DMaps$OVEN, breaks=breaks_qt$brks, col = topo.colors(7), colNA="grey70")
breaks_qt <- classInt::classIntervals(na.omit(getValues(for2DAndNf1DRes$OVEN)), n =7, style = "quantile")
plot(for2DAndNf1DRes$OVEN, breaks=breaks_qt$brks, col = topo.colors(7), colNA="grey70")

```

```{r plot maps for comparison}

getPlot1Dvs2D <- function(birdName,
                          phbRasterListName1D, 
                          phbRasterListName2D) {

  # NM <- eval(parse(text=paste("birdRasters$", birdName, sep = "")))
  ras1D <- eval(parse(text=paste(phbRasterListName1D, "$", birdName, sep = "")))
  # ras1D <- overlay(NM, ras1D, fun=function(x,y){return(x-y)})
  ras2D <- eval(parse(text=paste(phbRasterListName2D, "$", birdName, sep = "")))
  #ras2D <- overlay(NM, ras2D, fun=function(x,y){return(x-y)})
 names(ras1D) <- "Residuals post-hoc binning 1D"
 names(ras2D) <- "Residuals post-hoc binning 2D"


compRas <- raster::stack(ras1D, ras2D)


rasComparison <- gplot(compRas) + 
                  geom_tile(aes(fill = value)) +
                    facet_wrap(~ variable) +
                    scale_fill_gradientn(colours = rev(terrain.colors(500)), na.value = "grey70") +
  coord_equal() + 
    ggtitle(paste0(birdName)) +
     theme_classic() +
    theme(legend.position = "bottom",
        line = element_blank(),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        legend.key.width = unit(2, 'cm')) 
  
  
  return(rasComparison)
}

# plot1Dvs2DBBWO <- getPlot1Dvs2D(birdName = "BBWO", 
#                             phbRasterListName1D = "resForest1D", 
#                             phbRasterListName2D = "resForest2D")
# plot1Dvs2DCAWA <- getPlot1Dvs2D(birdName = "CAWA", 
#                             phbRasterListName1D = "resForest1D", 
#                             phbRasterListName2D = "resForest2D")
# 
plot1Dvs2D <- getPlot1Dvs2D(birdName = "BARS",
                            phbRasterListName1D = "for1DRes",
                            phbRasterListName2D = "for2DRes")
plot1Dvs2D
# plot1Dvs2DRUBL <- getPlot1Dvs2D(birdName = "RUBL", 
#                             phbRasterListName1D = "resForest1D", 
#                             phbRasterListName2D = "resForest2D")

#compareResRas <- gridExtra::grid.arrange(plot1Dvs2DBBWO, plot1Dvs2DCAWA, plot1Dvs2DOVEN, plot1Dvs2DRUBL) 
```




# ANALYSIS


```{r getResidualsHistogram}

resTabs <- lapply(X = birdList, FUN = function(bird){

   ras1D <- eval(parse(text=paste("for1DRes$", bird, sep = "")))
   ras2D <- eval(parse(text=paste("for2DRes$", bird, sep = "")))


  
  resVals1D <- data.table(getValues(ras1D))
  resVals1D <- setnames( resVals1D,  "resVals")
  resVals1D <- na.omit(resVals1D)
  res1DLab <- rep("res1D", nrow(resVals1D))
  resVals1D <- cbind(resVals1D, binningType = res1DLab)
  
  resVals2D <- data.table(getValues(ras2D))
  resVals2D <- setnames(resVals2D,  "resVals")
  resVals2D <- na.omit(resVals2D)
  res2DLab <- rep("res2D", nrow(resVals2D))
  resVals2D <- cbind(resVals2D, binningType = res2DLab)
  
  resVals <- rbind(resVals1D, resVals2D)
  birdSp <- rep(paste(bird), nrow(resVals))
  resVals <- cbind(resVals, birdSp = birdSp)
  print(resVals)
 
   return(resVals)
})
names(resTabs) <- birdList
resTabs

#as Rdata file

   save(resTabs,
        file =  file.path(outputFolderBirdPreds, "resTabs.Rdata"))
   # load(file.path(outputFolderBirdPreds, "resTabs.Rdata"))

resTabSingle <- rbind(resTabs$BARS, resTabs$GRAJ, resTabs$OVEN, resTabs$RCKI)

#as Rdata file

   save(resTabSingle,
        file =  file.path(outputFolderBirdPreds, "resTabSingle.Rdata"))
   # load(file.path(outputFolderBirdPreds, "resTabSingle.Rdata"))

# New facet label names for forest classes
binningType.labs <- c("1D Binning", "2D Binning")
names(binningType.labs) <- c("res1D", "res2D")
   
# make histogram of ages
  resHist <-ggplot(data = resTabSingle, aes(x= resVals, fill=factor(binningType))) + 
    geom_histogram() +
    ggtitle("Histograms of residuals in forest raster areas") +
     xlab("Residual Value")  +
    ylab("Pixel Count") +
    facet_grid( binningType ~ birdSp, 
    labeller = labeller(binningType = binningType.labs), scales="free") +
    scale_fill_manual(values=c('skyblue1', "springgreen4")) + 
    theme_classic() +
    theme(title = element_text(size = 14),
      strip.text.x = element_text(
        size = 11.5),
      strip.text.y = element_text(
        size = 11.5),
      legend.position = "NULL",
      axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
      axis.title = element_text(size = 12)) 
  
  resHist
  





```


```{r get mean and spatial autocorrelation of residual data}


residualStats <- lapply(X = birdList, FUN = function(bird){
  
  
  res1D <- eval(parse(text=paste("for1DRes$", bird, sep = "")))
  res2D <- eval(parse(text=paste("for2DRes$", bird, sep = "")))
 

m1D <- mean(res1D[], na.rm = TRUE)
m2D <- mean(res2D[], na.rm = TRUE)
sa1D  <- raster::Moran(res1D)
sa2D  <- raster::Moran(res2D)

residualStats <- matrix(c( m1D, m2D,  sa1D, sa2D), ncol= 4, byrow=TRUE)
colnames(residualStats) <- c( 'mean1D', 'mean2D', "spaAutocor1D", "spaAutocor2D")
row.names(residualStats) <- bird
print(paste(bird, " calculation complete"))
return(residualStats)
  
    })

residualStats <- do.call(rbind, residualStats)

residualStats

 save(residualStats,
        file =  file.path(outputFolderBirdPreds, "residualStats.Rdata"))
   # load(file.path(outputFolderBirdPreds, "residualStats.Rdata"))

head(residualStats)



```


```{r plot means and spatial autocorrelation}


   tab <- melt(residualStats[,1:2])
   colnames(tab) <- c( "birdSp", "binType", "meanResidual")
   tab


plotMeanRes <- ggplot(data = tab, 
                                aes(fill= binType, y= meanResidual, x= birdSp)) + 
    geom_bar(position = "dodge", stat = "identity")+
    ggtitle("Mean values of forest area residual rasters, by species") +
    xlab("Species")  +
    ylab("Mean") +
    theme_classic() + 
    scale_fill_manual(values=c('skyblue1', "springgreen4"), 
                      labels = c("1D", "2D")) +
    theme(title = element_text(size = 14),
          legend.title = element_blank(),
          legend.position="bottom",
          axis.text.x = element_text(size = 10, angle = 45, hjust = 1), 
          axis.title = element_text(size = 12))
        


plotMeanRes

    
 
 


# SPATIAL AUTOCORRELATION

   tabSA <- melt(residualStats[,3:4])
   colnames(tabSA) <- c( "birdSp", "binType", "spatialAutocorrelation")
   tabSA


plotSARes <- ggplot(data = tabSA, 
                                aes(fill= binType, y= spatialAutocorrelation, x= birdSp)) + 
    geom_bar(position = "dodge", stat = "identity") +
    ggtitle("Moran test for spatial autocorrelation results for forest area residual rasters, by species") +
    xlab("Species")  +
    ylab("Spatial Autocorrelation") +
    theme_classic() + 
    scale_fill_manual(values=c('skyblue1', "springgreen4"), 
                      labels = c("1D", "2D")) +
    theme(title = element_text(size = 14),
          legend.title = element_blank(),
          legend.position= "bottom",
          axis.text.x = element_text(size = 10, angle = 45, hjust = 1), 
          axis.title = element_text(size = 12))
plotSARes

    
  

```


```{r do spearman rank correlation test on mapped bins}

spearmanStats <- lapply(X = birdList, FUN = function(bird){
  
  print(bird) 
  
  nmRas <- eval(parse(text=paste("birdRasters$", bird, sep = "")))
   map1D <- eval(parse(text=paste("for1DMaps$", bird, sep = "")))
   map2D <- eval(parse(text=paste("for2DMaps$", bird, sep = "")))

  
valsNM <- getValues(nmRas)
vals1DMap <- getValues(map1D)
vals2DMap <- getValues(map2D)
valsMaps <- cbind(valsNM, vals1DMap, vals2DMap) 
valsMaps <- na.omit(valsMaps)
valsMaps <- as.data.table(valsMaps)
head(valsMaps)
#Check normality assumption 
#Shapiro-Wilk normality test for all data
# ad.test(valsMaps$valsNM) # => p = 0.1229
# ad.test(valsMaps$vals1DMap) # => p = 0.09
# ad.test(valsMaps$vals2DMap)

# library("ggpubr")
# ggqqplot(valsMaps$valsNM, ylab = "National Model Prediction")
# ggqqplot(valsMaps$vals1DMap, ylab = "1D Map Prediction")
# ggqqplot(valsMaps$vals2DMap, ylab = "2D Map Prediction")

spearman1D <-cor(valsMaps$valsNM, valsMaps$vals1DMap,  method = "spearman")
spearman1D
spearman2D <-cor(valsMaps$valsNM, valsMaps$vals2DMap,  method = "spearman")
spearman2D



spearmanStats <- matrix(c( spearman1D, spearman2D), ncol= 2, byrow=TRUE)
colnames(spearmanStats) <- c( 'spearman1D', 'spearman2D')
row.names(spearmanStats) <- bird

return(spearmanStats)

    })

spearmanStats <- do.call(rbind, spearmanStats)

spearmanStats

 save(spearmanStats,
        file =  file.path(outputFolderBirdPreds, "spearmanStats.Rdata"))
   # load(file.path(outputFolderBirdPreds, "spearmanStats.Rdata"))


   tabST <- melt(spearmanStats[,1:2])
   colnames(tabST) <- c( "birdSp", "binType", "spearmanTestStat")
   tabST


plotSTRes <- ggplot(data = tabST, 
                                aes(fill= binType, y= spearmanTestStat, x= birdSp)) + 
    geom_bar(position = "dodge", stat = "identity") +
    ggtitle("Spearman rank correlation test for national models with forest area predictions, by species") +
    xlab("Species")  +
    ylab("Spearman test statistic") +
    theme_classic() + 
    scale_fill_manual(values=c('skyblue1', "springgreen4"), 
                      labels = c("1D", "2D")) +
    theme(title = element_text(size = 14),
          legend.title = element_blank(),
          legend.position= "bottom",
          axis.text.x = element_text(size = 10, angle = 45, hjust = 1), 
          axis.title = element_text(size = 12))
plotSTRes


```


```{r do paired t-test on mean residuals and spatial autocorrelation of the residuals results}
residualStatsTTest <- as.data.frame(residualStats)
 
#mean residual
diffMeansRes <- residualStatsTTest$mean1D - residualStatsTTest$mean2D
hist(diffMeansRes,   
     col="darkblue", 
     main="Histogram of differences between mean residuals for 1D and 2D by bird",
     xlab="Difference")
meanResTTest <- t.test(residualStatsTTest$mean1D,
       residualStatsTTest$mean2D,
       paired=TRUE)
meanResTTest 

#spatial autocorrelation of the residuals
diffSpaAutocor <- residualStatsTTest$spaAutocor1D - residualStatsTTest$spaAutocor2D
hist(diffSpaAutocor,   
     col="darkblue", 
     main="Histogram of differences between tests of spatial autocorrelation of the residuals for 1D and 2D by bird",
     xlab="Difference")
spaAutoCorTTest <- t.test(residualStatsTTest$spaAutocor1D,
       residualStatsTTest$spaAutocor2D,
       paired=TRUE)
spaAutoCorTTest 

```